---
title: "Linear Model Assumptions"
author: "Marc Los Huertos"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Anscombe Data

The first thing I did is create a dataframe with all the data, where a matrix() is converted to a dataframe. See below.

Note: I removed the label 'cars' because it wasn't useful to keep.

```{r}
Anscombe = as.data.frame(
  matrix(c(
    1, 10, 10, 10, 8, 8.04, 9.14, 7.46, 6.58, 
    2, 8, 8, 8, 8, 6.95, 8.14, 6.77, 5.76,
    3, 13, 13, 13, 8, 7.58, 8.74, 12.74, 7.71,
    4,  9, 9, 9, 8, 8.81, 8.77, 7.11, 8.84,
    5, 11, 11, 11, 8, 8.33, 9.26, 7.81, 8.47,
    6, 14, 14, 14, 8, 9.96, 8.10, 8.84, 7.04,
    7,  6, 6, 6, 8, 7.24, 6.13, 6.08, 5.25,
    8,  4, 4, 4, 19, 4.26, 3.10, 5.39, 12.50,
    9, 12, 12, 12, 8, 10.84, 9.13, 8.15, 5.56,
    10, 7, 7, 7, 8, 4.82, 7.26, 6.42, 7.91,
    11, 5, 5, 5, 8, 5.68, 4.74, 5.73, 6.89), 
    nrow=11,  byrow=TRUE))
```

After creating the data frame, I renamed the variables using names().

```{r}
names(Anscombe) = c("Obs.", "x1", "x2", "x3", "x4", "y1", "y2", "y3", "y4")
```


Finally, we can print out the dataframe with just the name of the dataframe.

```{r}
Anscombe
```


## Plot Each Dataset with the Dataframe

Below I plot each dataset into a separate graph.

```{r, echo=TRUE}
plot(y1 ~ x1, Anscombe, main="Dataset A")
plot(y2 ~ x2, Anscombe, main="Dataset B")
plot(y3 ~ x3, Anscombe, main="Dataset C")
plot(y4 ~ x4, Anscombe, main="Dataset D") 
```

## Plots with Best Fit Lines in a Panel

So, first I'll add a command to make a panel so we can see all the graphs at once. Pretty conveneient. 

Next, I createa linear model, using lm(), where we want the y is a function of x (y~x). 

Finally, I added a line, abline(), with the coeffients from the linear model, coef().

```{r}
par(mfrow=c(2,2))
plot(y1 ~ x1, Anscombe, main="Dataset A")
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm)); 

plot(y2 ~ x2, Anscombe, main="Dataset B")
  B.lm = lm(y2 ~ x2, Anscombe); abline(coef(B.lm))

plot(y3 ~ x3, Anscombe, main="Dataset C")
  C.lm = lm(y3 ~ x3, Anscombe); abline(coef(C.lm))

plot(y4 ~ x4, Anscombe, main="Dataset D")
  D.lm = lm(y4 ~ x4, Anscombe); abline(coef(D.lm))
```

## Changing the scales so each graph matches

First, I had to determine the minimum and maximum for the Xs and Ys, where plot(y~x, dataframe, xlim=c(min, max). 

We need to get separate min for the Xs and Ys and max for the Xs and Ys. 

It's a bit tricky because I had to select several columns...but not the first!

```{r}
names(Anscombe)
```

So, we need columns, 2-5 for the Xs and 6-9 for the Ys. for example, here is the minimum for the Xs. 

```{r}
min(Anscombe[, 2:5])
```

So, instead of put this value in each plot command, we can put the to values (min and max) as a vector for each plot function used.

As a trick, it saves sometime, but works when you are a bit further in R, so I'll show you both ways!


### Manually defining max X and min X and max Y and min Y.

I define a new vector with the mins and maxs. NOTE: I can put to commands on the same line with a semicolon.

```{r}
xmin = min(Anscombe[, 2:5]); xmin
xmax = max(Anscombe[, 2:5]); xmax

ymin = min(Anscombe[, 6:9]); ymin
ymax = max(Anscombe[, 6:9]); ymax
```

And now the plots. Notice how each plot has the min and max.
```{r}
par(mfrow=c(2,2), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=c(4,19), ylim=c(3.1,12.74));
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm))

plot(y2 ~ x2, Anscombe, main="Dataset B", xlim=c(4,19), ylim=c(3.1,12.74)) 
  B.lm = lm(y2 ~ x2, Anscombe); abline(coef(B.lm))

plot(y3 ~ x3, Anscombe, main="Dataset C", xlim=c(4,19), ylim=c(3.1,12.74)) 
  C.lm = lm(y3 ~ x3, Anscombe); abline(coef(C.lm))

plot(y4 ~ x4, Anscombe, main="Dataset D", xlim=c(4,19), ylim=c(3.1,12.74)) 
  D.lm = lm(y4 ~ x4, Anscombe); abline(coef(D.lm))
```

I don't like this method because it relies on me typing the right number -- I'd rather define the numbers and then use a vector that contains both the max and min and defined here:

```{r}
xminmax = c(xmin, xmax); xminmax
yminmax = c(ymin, ymax); yminmax
```

Now let's do the plots, same results!  
```{r}
par(mfrow=c(2,2), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=xminmax, ylim=yminmax);
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm))

plot(y2 ~ x2, Anscombe, main="Dataset B", xlim=xminmax, ylim=yminmax) 
  B.lm = lm(y2 ~ x2, Anscombe); abline(coef(B.lm))

plot(y3 ~ x3, Anscombe, main="Dataset C", xlim=xminmax, ylim=yminmax) 
  C.lm = lm(y3 ~ x3, Anscombe); abline(coef(C.lm))

plot(y4 ~ x4, Anscombe, main="Dataset D", xlim=xminmax, ylim=yminmax)
D.lm = lm(y4 ~ x4, Anscombe); abline(coef(D.lm))
```

## Adding Text with the coeficients

Finally, we can add text to the plots using text() and paste() text with a R function, e.g. coef(). I also rounded the values to 3 digits and made the best fit line blue.

NOTE: coef() has two values -- slope and intercept!

```{r}
coef(A.lm)
```

We can extract them individually using square brackets

```{r}
coef(A.lm)[1]
coef(A.lm)[2]
```

where position 1 is for the intercept and position 2 is the slope.


```{r}
par(mfrow=c(2,2), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=xminmax, ylim=yminmax); 
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm), col="blue"); 
  text(14,5, paste("slope =", round(coef(A.lm)[2], 3)));
  text(14,4, paste("intercept =", round(coef(A.lm)[1], 3)))

plot(y2 ~ x2, Anscombe, main="Dataset B", xlim=xminmax, ylim=yminmax); 
  B.lm = lm(y2 ~ x2, Anscombe); abline(coef(B.lm), col="blue")
  text(14,6, paste("slope =", round(coef(B.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(B.lm)[1], 3)))

plot(y3 ~ x3, Anscombe, main="Dataset C", xlim=xminmax, ylim=yminmax); 
  C.lm = lm(y3 ~ x3, Anscombe); abline(coef(C.lm), col="blue")
  text(14,6, paste("slope =", round(coef(C.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(C.lm)[1], 3)))

  plot(y4 ~ x4, Anscombe, main="Dataset D", xlim=xminmax, ylim=yminmax); 
  D.lm = lm(y4 ~ x4, Anscombe); abline(coef(D.lm), col="blue")
  text(14,6, paste("slope =", round(coef(D.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(D.lm)[1], 3)))

```

Can you think of things to make this a better graph? Slack suggestions and I'll add it!

## Statistics

Are the results signficant?  Good question, we can use summary() to get the results:

```{r}
summary(A.lm)
summary(B.lm)
summary(C.lm)
summary(D.lm)
```

From this we can determine if the model is outside the range of random variation. 

I each case the p-values are <0.05, suggesting significant. In addition, we can look at the R2, which is an estimate of how much of the variation of y is explained by x. In nearly every case these value are above .62, which is generally thought of as quite good as a predictor. Not perfect, which would be 1, but not close to zero, where there would be not explanatory value of x. 

## What's going on!

We need to look at the assumptions, which relies on plots() of the linear model -- which generates 4 default plots. 

```{r}
par(mfrow=c(2,2)); plot(A.lm); par(mfrow=c(1,1))
```

Let's compare each model with their assumptions. There should be no structure in first diagnostic plot!

There isn't much that stands out for the others.

```{r}
par(mfrow=c(2,5), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=xminmax, ylim=yminmax); 
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm), col="blue"); 
  text(14,5, paste("slope =", round(coef(A.lm)[2], 3)));
  text(14,4, paste("intercept =", round(coef(A.lm)[1], 3)))

plot(A.lm)

plot(y2 ~ x2, Anscombe, main="Dataset B", xlim=xminmax, ylim=yminmax); 
  B.lm = lm(y2 ~ x2, Anscombe); abline(coef(B.lm), col="blue")
  text(14,6, paste("slope =", round(coef(B.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(B.lm)[1], 3)))

plot(B.lm)
```

```{r}
par(mfrow=c(2,5), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=xminmax, ylim=yminmax); 
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm), col="blue"); 
  text(14,5, paste("slope =", round(coef(A.lm)[2], 3)));
  text(14,4, paste("intercept =", round(coef(A.lm)[1], 3)))

plot(A.lm)

plot(y3 ~ x3, Anscombe, main="Dataset C", xlim=xminmax, ylim=yminmax); 
  C.lm = lm(y3 ~ x3, Anscombe); abline(coef(C.lm), col="blue")
  text(14,6, paste("slope =", round(coef(C.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(C.lm)[1], 3)))

plot(C.lm)
```


```{r}
par(mfrow=c(2,5), las=1)
plot(y1 ~ x1, Anscombe, main="Dataset A", xlim=xminmax, ylim=yminmax); 
  A.lm = lm(y1 ~ x1, Anscombe); abline(coef(A.lm), col="blue"); 
  text(14,5, paste("slope =", round(coef(A.lm)[2], 3)));
  text(14,4, paste("intercept =", round(coef(A.lm)[1], 3)))

plot(A.lm)

plot(y4 ~ x4, Anscombe, main="Dataset D", xlim=xminmax, ylim=yminmax); 
  D.lm = lm(y4 ~ x4, Anscombe); abline(coef(D.lm), col="blue")
  text(14,6, paste("slope =", round(coef(D.lm)[2], 3)));
  text(14,5, paste("intercept =", round(coef(D.lm)[1], 3)))

plot(D.lm)
```

