---
title: "Empathy or Something Else: H. R. 6637"
author: "Marc Los Huertos"
date: "4/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How do Americans think about immigration?

Americans are taught to fear immigration. We are told that immigrants compete with our jobs. We are told immigrants take advantage of our public services. We are told that immigrants are a threat to our American way of life (language, religion, and whiteness). 

At the same time, many Americans are immigrants and proud of their immigrant history. Of course, some migration was culmpulstory, i.e. as part of the institution of slavery and the slave trade. And of course, the indigenous were displaced and had their lands stolen as part of the America project. 

It's no wonder Americans are confused in how they might or should think about immigration history and immigration history. 

## How can we be inclusive while being insecure about job security in the face of immigration?

The current standard for evaluating policies at the border is not whether the policies are effective or consistent with our values but whether they are “tough.” It is as if we judged college and high school basketball coaches based on who yelled at and berated their players the most.

Ali Noorani points out that a great deal of media reporting, particularly on television, focuses on the border, but that Americans see or read little about what people in Mexico and Central America leave behind. He believes a better understanding of the violence and economic desperation people experience might elicit more sympathy and a greater willingness to craft political compromises on immigration. It’s a sensible conclusion.

## Is immigration scape-goating?

```{r cars}
summary(cars)
```

## What has Ken Calvert said and done?

## What does Ken think or do about Comprehensive Immigration Reform?

### NGOs and Think Tanks

Just as the politics of immigration policy is highly polarized, I suspect the policy think tanks, organizations, and advocacy groups are similarly polarized. Thus, that there might be a unbiased, objective view is probably a fantasy. Nevertheless, let's see what I can learn in an afternoon and then communicated it to the readers. 

1. Migration Policy Institute

Mission: The nonpartisan Migration Policy Institute seeks to improve immigration and integration policies through authoritative research and analysis, opportunities for learning and dialogue, and the development of new ideas to address complex policy questions.

<https://www.migrationpolicy.org/topics/comprehensive-immigration-reform>

2. American Immigration Council

Our Vision : The Council believes that immigrants are part of our national fabric, bringing energy and skills that benefit all Americans. We envision an America that values fairness and justice for immigrants and advances a prosperous future for all. 

<https://www.americanimmigrationcouncil.org/>



3. Pew Research Center 

Pew Research Center is a nonpartisan fact tank that informs the public about the issues, attitudes and trends shaping the world. We conduct public opinion polling, demographic research, content analysis and other data-driven social science research. We do not take policy positions.

MISSION
The Center generates a foundation of facts that enriches the public dialogue and supports sound decision-making. Our empirical research on a wide range of topics helps U.S. and international policymakers, civic leaders, educators and the public at large understand and address some of the world’s most challenging problems. Our public opinion surveys allow the voice of the people to be heard, and our demographic, economic and political analyses provide context to understand how the world is changing. We are nonprofit, nonpartisan and nonadvocacy. Our mission is to inform, not to prescribe. We believe that better information can build a better world. “Fact-based information is the fuel democracies run on—the raw material from which societies identify problems and construct solutions.”

<https://www.pewresearch.org/>

4. Council on Foreign Relations

The Council on Foreign Relations (CFR) is an independent, nonpartisan membership organization, think tank, and publisher dedicated to being a resource for its members, government officials, business executives, journalists, educators and students, civic and religious leaders, and other interested citizens in order to help them better understand the world and the foreign policy choices facing the United States and other countries. Founded in 1921, CFR takes no institutional positions on matters of policy. Our goal is to start a conversation in this country about the need for Americans to better understand the world.

<https://www.cfr.org/>

5. FAIR

Our Objectives

FAIR seeks to reduce overall immigration to a more normal level. Reducing legal immigration levels from well over one million at present to a very generous 300,000 a year over a sustained period will allow America to manage growth, address environmental concerns, and maintain a high quality of life.

FAIR puts the interests of American citizens and future generations ahead of big business and partisan demands.

<https://www.fairus.org/>

6. Heritage Foundation

Heritage’s mission is to formulate and promote public policies based on the principles of free enterprise, limited government, individual freedom, traditional American values, and a strong national defense.

<https://www.heritage.org/>

# Citations

Bureau of Labor Statistics, U.S. Department of Labor, The Economics Daily, Foreign-born workers made up 17.4 percent of labor force in 2019 at https://www.bls.gov/opub/ted/2020/foreign-born-workers-made-up-17-point-4-percent-of-labor-force-in-2019.htm (visited April 04, 2022).
