---
title: "Empathy or Something Else: H. R. 6637"
author: "Marc Los Huertos"
date: "4/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How do Americans think about immigration?

Americans are taught to fear immigration. We are told that immigrants compete with our jobs. We are told immigrants take advantage of our public services. We are told that immigrants are a threat to our American way of life (language, religion, and whiteness). 

The current standard for evaluating policies at the border is not whether the policies are effective or consistent with our values but whether they are “tough.” It is as if we judged college and high school basketball coaches based on who yelled at and berated their players the most.

Ali Noorani points out that a great deal of media reporting, particularly on television, focuses on the border, but that Americans see or read little about what people in Mexico and Central America leave behind. He believes a better understanding of the violence and economic desperation people experience might elicit more sympathy and a greater willingness to craft political compromises on immigration. It’s a sensible conclusion.

## Is immigration scape-goating?

```{r cars}
summary(cars)
```

## What has Ken Calvert said and done?

## What does Ken think or do about Comprehensive Immigration Reform?

### NGOs and Think Tanks

Migration Policy Institute

<https://www.migrationpolicy.org/topics/comprehensive-immigration-reform>
