---
title: "Calvert and Climate Science"
author: "Marc Los Huertos"
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Opps, Wrong Side of History

I actually don't like the climate denier or skeptic terminology. It's not that I don't like the derogatory nature or that I don't like doing public policy by framing an issue as a knowledge gap. 

No, instead, I want to focus on the public good. What is best for Americans, and for others around the world?

## A stable climate to thrive
You can also embed plots, for example:

```{r pressure, echo=FALSE}
calvert.csv <- "/home/CAMPUS/mwl04747/github/beginnersluck/CD41/ken-s. calvert-export (1).csv"
cc <- read.csv(calvert.csv)
cc$YN[cc$Vote=="-"]<- 0
cc$YN[cc$Vote=="?"]<- NA
cc$YN[cc$Vote=="+"]<- 1
str(cc)
```

```{r}
Total = aggregate(cc$YN, by=list(cc$YN, cc$Year), length)
calvert = aggregate(cc$YN, by=list(cc$YN, cc$Year), sum)
Table = data.frame(Year = Total$Group.2, Bills = Total$x,
                   Yes = calvert$x)
Table
```
