---
title: "Untitled"
author: "Marc Los Huertos"
date: "2022-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(deSolve)
library(ggplot2)
library(gridExtra)
library(grid)
library(dplyr)
```

# Define Scope

An upper limit exists, but it's not worth modeling

Carbon Dioxode Concentration have no risen linearly

Good records of CO2 in atmosphere

Weaker records on inputs/outputs

# Start 1890
# End present day
# Model CO2 as a contentration

# Emitted CO2 is then diluted and added to atmsophere

```{r}

# Volume of Atmosphere 

v = 4200000000 # km3

# Using Mauna Loa to estimate Total CO2 Stock
ppmv = 420 # mg/L or place holder until I get ppm_t sorted

GtC = ppmv * 2.3; GtCO2 <- GtC *44/12; GtCO2




km3perL = 1e12
Gtperkg = 1e12
mgperkg = 1e6

mgperkm3 = 420*1e12
Gtperkm3 = mgperkm3/(mgperkg*Gtperkg); Gtperkm3

(AtmCO2 = Gtperkm3 * v); 

Gt/420

v * .04

7.82*420

# Removal is "constant rate, but lower than emissions

```


The Bern Model!, let's try this... is a fraction of the air left in the atmosphere after time, t, it was released. approximatley 21% never leaves.

$$f(t)=0.217+0.338^{(-t/18.51)}+0.259^{(-t/172.9)} +0.186^{(-t/1.186)}$$
```{r}

t = seq(1, 100, by=1); t

bern <- function(t){
  0.217 + 0.338^(-t/18.51)+0.259^(-t/172.9) +0.186^(-t/1.186)
}

bern(t)

```


## Stocks and flow model...


```{r}
# Set the time period and step. Define the stocks and auxiliaries.
START <- 0
FINISH <- 100
STEP <- 0.25

start = 1750
finish = 2023
step = 1
simtime = seq(start, finish, by=step); simtime

simtime <- seq(START, FINISH, by = STEP)
stocks <- c(sStock = 100)
auxs <- c(aCapacity = 10000, 
          aRef.Availability = 1,
          aRef.GrowthRate = 0.10)

```


## Annual Emissions

```{r}
emissions = read.csv("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv")
str(emissions); nrow(emissions)
unique(emissions$country)

emissions$iso_code[emissions$country=="Upper-middle-income countries"]

sum(!is.na(emissions$iso_code))

country_emissions = subset(emissions, subset=(iso_code!="")); nrow(country_emissions)

yearlyemissions = country_emissions %>%
  group_by(year) %>%
  summarise(
    emissions = sum(co2, na.rm=T)
  )

plot(emissions/1000 ~ year, data=yearlyemissions, ylab="Billion Tons")

baseline = 280
baselineGt = 280*v/



```

```{r}

# Create the function
model <- function(time, stocks, auxs){
  with(as.list(c(stocks, auxs)),{
    aAvailability <- 1 - sStock / aCapacity
    aEffect <- aAvailability / aRef.Availability
    aGrowth.Rate <- aRef.GrowthRate * aEffect
    fNet.Flow <- sStock * aGrowth.Rate
    dS_dt <- fNet.Flow
    return(list(c(dS_dt), NetFlow = fNet.Flow,
                GrowthRate = aGrowth.Rate, 
                Effect = aEffect,
                Availability = aAvailability))
  })
}

# Create the data frame
sModel <- data.frame(ode(y = stocks, times = simtime, func = model,
                         parms = auxs, method = "euler"))

head(sModel)

# Run the following snippet to enable `multiplot`
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

# Plot the outputs using `multiplot`
stockTimePlot <- sModel %>% 
  ggplot() +
  geom_line(aes(time, sStock), color = "purple")

availTimePlot <- sModel %>% 
  ggplot() +
  geom_line(aes(time, Availability), color = "blue")

netflowTimePlot <- sModel %>% 
  ggplot() +
  geom_line(aes(time, NetFlow), color = "red")

growthrateTimePlot <- sModel %>% 
  ggplot() +
  geom_line(aes(time, GrowthRate), color = "green")

multiplot(stockTimePlot, netflowTimePlot, availTimePlot, growthrateTimePlot, cols = 2)
