---
title: "Understanding Cap and Trade Systems: An Interactive Simulation Guide"
author: "Environmental Economics Laboratory"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", out.width = "80%")
library(ggplot2)
```

## Abstract

This guide introduces students to tradable emission permits through an interactive simulation. You will learn the history, theory, and practical application of cap-and-trade systems—one of the most powerful market-based tools for environmental regulation.

---

## The Origins of Tradable Permits

The concept of tradable pollution permits emerged from economic theory in the 1960s.\
[^dales1968]: Dales, J.H. (1968). *Pollution, Property & Prices*. University of Toronto Press.

John Dales first proposed that pollution rights could be bought and sold like any other commodity, creating incentives for efficient pollution reduction.

### Why Tradable Permits?

Traditional environmental regulation often uses *command-and-control* approaches[^cac]: telling each polluter exactly how much to reduce.

[^cac]: Command-and-control regulations specify exactly how much each polluter must reduce emissions, regardless of the cost.

This is inefficient because:

- Different firms face different costs of reducing pollution  
- A firm that can reduce cheaply should reduce more  
- A firm facing high reduction costs should reduce less  
- Markets can find this efficient allocation automatically  

> **Key Insight:** Let those who can reduce pollution cheaply do more of the reduction, while those facing high costs can buy permits instead.

The tradable permit system harnesses market forces to achieve environmental goals at the lowest possible cost to society.

---

## Real-World Applications

### The U.S. Acid Rain Program (1990)

The most celebrated success story is the U.S. SO₂ trading program under the Clean Air Act Amendments of 1990.\
[^so2]: Schmalensee, R., & Stavins, R.N. (2013). *The SO₂ Allowance Trading System: The Ironic History of a Grand Policy Experiment.* *Journal of Economic Perspectives*, 27(1), 103–122.

- **Goal:** Reduce sulfur dioxide emissions by 50%  
- **Result:** Goal achieved two years early  
- **Cost:** One-third of initial estimates  
- **Innovation:** Spurred development of scrubber technology  

### The European Union Emissions Trading System

Launched in 2005, the EU ETS is the world’s largest carbon market, covering over 11,000 installations.\
[^euets]: As of 2024, the EU ETS covers approximately 40% of the EU's greenhouse gas emissions.

### California’s Cap-and-Trade Program

California’s program, started in 2013, covers about 80% of the state’s greenhouse gas emissions and has been linked with Quebec’s system since 2014.

---

## Interactive Simulation

This section provides an R-based simulation of a simplified cap-and-trade market with heterogeneous firms.

```{r simulate-market}
simulate_market <- function(n_firms = 24, cap = 100, price_floor = 20, price_ceiling = 150) {
  set.seed(123)
  cost_level <- runif(n_firms, 0.5, 1.5)
  wtp <- price_floor + (price_ceiling - price_floor) * cost_level
  allocation <- rep(cap / n_firms, n_firms)
  desired <- runif(n_firms, 0, cap / n_firms * 1.5)
  surplus <- allocation - desired
  data.frame(Firm = 1:n_firms, CostLevel = cost_level, WTP = wtp,
             Allocation = allocation, Desired = desired, Surplus = surplus)
}

market_data <- simulate_market()
head(market_data)
```

### Supply and Demand Curves

```{r market-curves, fig.height=5}
ggplot(market_data, aes(x = WTP)) +
  geom_histogram(fill="gray80", color="black", bins=15) +
  labs(title="Distribution of Willingness to Pay for Permits",
       x="Willingness to Pay ($)", y="Number of Firms")
```

This histogram approximates the demand for permits in the market.  
We can also visualize stylized supply and demand curves.

```{r supply-demand, fig.height=5}
price <- seq(20, 150, length.out = 50)
demand <- 120 - 0.7 * (price - 20)
supply <- 0.6 * (price - 20)
df <- data.frame(price, demand, supply)

ggplot(df, aes(x=price)) +
  geom_line(aes(y=demand, color="Demand")) +
  geom_line(aes(y=supply, color="Supply")) +
  geom_vline(xintercept=80, linetype="dashed") +
  annotate("text", x=83, y=60, label="Equilibrium", hjust=0) +
  labs(title="Stylized Cap-and-Trade Market Equilibrium",
       x="Permit Price ($)", y="Quantity (tons)",
       color="Curve") +
  theme_minimal()
```

---

## Key Economic Concepts

### Equimarginal Principle

Tradable permits achieve the *equimarginal principle*[^equimarginal]: all firms face the same marginal abatement cost (the permit price).

[^equimarginal]: At the optimum, the marginal cost of abatement is equal across all firms.

- Low-cost firms reduce and sell  
- High-cost firms buy instead of reducing  
- The cap is achieved at minimum total cost  

### Static vs. Dynamic Efficiency

- **Static efficiency:** Achieve today’s target at lowest cost  
- **Dynamic efficiency:** Encourage innovation for the future  

---

## Conclusion

Cap-and-trade systems harness private incentives to achieve public environmental goals efficiently. The simulation helps you understand firm behavior, market dynamics, and the efficiency of market-based environmental regulation.

**Reflection:** The real challenge lies in balancing stringency vs. cost, efficiency vs. equity, and simplicity vs. precision—just as real-world policymakers must.
