---
title: "Pb and Parks"
author: "Marc Los Huertos"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

As we have learned, R can be powerful, but it's often hard to remember what to do if it's been a while since you have used it. I have created this document to help you analyze the Pb data -- why?  Because I am more interested in having you understand what the analysis process is than the actual R code. 

But some understanding of R is useful so you can see the steps needed for the analysis!

## Read the Data

First, we read the data. I like to create a filename object that I can modify based on my own path and filename. 


```{r readcsv}
filename = "/home/CAMPUS/mwl04747/github/beginnersluck/Pb_and_Parks/EA 30 Data Entry - Sheet1.csv"
import <- read.csv(filename)
str(import)
```

I fixed some of the values that were treated as factors, i.e. categorical data. For example, Pb has "<0.0001" which was treated as a character string instead of a numeric value -- this the entire variable was being treated as a character string. We fixed that by changing "<0.0001" to 0. I haven't done this for other variables yet. 

## Exploratory Analysis

We can easily make a histogram as below...
```{r exploratory}
hist(import$Pb208)
```

However, we only want to use the sample data, not the standards and blanks to calibrate the analysis run. See the Column A in the google sheet. So, let's subset the data to use only the sample data. 

```{r histsamples}
hist(import$Pb208[import$Type=="Sample"])

```

Let's clean this up a bit:

```{r histoclean}
hist(import$Pb208[import$Type=="Sample"], xlab="Pb Concentration (ppb)")

```

Based on this, let's create a dataset that only has the samples, so we don't have to worry about this for the rest of the analysis. To do this we can use the subset() function. 

```{r subset}
parkdata = subset(import, subset=import$Type=="Sample" & import$Sample.Name!="Blank")

```

## Converting the Values on a per gram of soil basis

Currently, we have a concentration based on 10 grams of soil that was digested and diluted upto 100 mL. So, we next need to change our concentration in the digestion relative to the soils (ug/g -> mg/kg)
```{r convertunits}
parkdata$SoilPb = parkdata$Pb208/10 
```

```{r boxplot}
boxplot(SoilPb ~ PARK, data=parkdata)
```

Let's fix the names of the parks and then adjust the size to fit on the boxplot.

```{r fixnames}
droplevels(parkdata$PARK)
levels(parkdata$PARK) <- c("Unknown", "Blaisdell", "CPark", "Jaeger", "La Puerta", "Larkin", "Marston", "Memorial", "R. Torrez", "Walker", "Wheeler")
```

```{r boxplot2}
par(las=1)
boxplot(SoilPb ~ PARK, data=parkdata, xlab="", xaxt="n")
axis(1, 1:11, parkdata$PARK, labels=F)

text(x = 1:length(parkdata$PARK),
     y = par("usr")[3] - 40,
     labels = unique(levels(parkdata$PARK)),
     xpd = NA,
     ## Rotate the labels by 35 degrees.
     srt = 35,
     cex = 1.2)
```

## Null Hypothesis Testing

No difference between Pb by parks.

```{r anova}
pb.aov <- aov(SoilPb ~ PARK, data=parkdata)
summary(pb.aov)
par(mfrow=c(2,2))
plot(pb.aov)
par(mfrow=c(1,1))
```

Testing distance

```{r distance}
pb.lm <- lm(Pb208 ~ DISTANCE66, data=parkdata)
summary(pb.lm)
par(mfrow=c(2,2))
plot(pb.lm)
par(mfrow=c(1,1))
plot(Pb208 ~ DISTANCE66, data=parkdata)
```






